Simulation study in rectangular domain

### 2024-05-02 R/simulation_study/simulation_study_fjords.Rmd
improve simulation study

# SET UP
## Import libraries and data

@@

@@

```{r}
#number of batch
N_BATCH=1

@@

## Estimate from high frequency data without error : ExpShore and tensor splines ti

### Use ExpShore covariate and tensor splines te

#### Fit sde

@@

crcvm_fjords_hf_ne1<- foreach(k=1:N_BATCH,.packages=c("smoothSDE")) %dopar% {
  #estimation with tensor splines
  formulas <- list(mu1=~1,mu2=~1,tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),omega=~te(theta,ExpShore,k=SP_DF,bs="cs"))
  par0 <- c(0,0,1,1,0)
  m <- SDE$new(formulas = formulas,data = data_fjords_hf_ne[[paste("BATCH",k,sep="")]],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2"),
                      other_data=list("H"=H_hf))
  m$fit(method="BFGS")
  m 
}

@@

#### Plot estimated smooth parameter omega in 3D for one batch

@@

#### Violin plot of the estimated spline coefficients for one batch

@@

#### Violin plot of the estimated spline coefficients for all batches

@@

for (k in 1:N_BATCH) {
  m<-crcvm_fjords_hf_ne1[[paste("BATCH",k,sep="")]]
  coeffs=m$coeff_re()
  coeff_names=rownames(coeffs)
  coeff_values=as.numeric(coeffs)
  sub_index=grepl("omega",coeff_names)
  sub_coeffs_df=data.frame("estimate"=coeff_values[sub_index],"coeff_name"=factor(coeff_names[sub_index]))
  coeffs_df=rbind(coeffs_df,sub_coeffs_df)
}

@@

p=ggplotly(p)
saveWidget(p, file ="crcvm_fjords_hf_ne1/all_batches_coeffs_crcvm_fjords_hf_ne1.png")

@@

#### Simulate from fitted model

@@

@@

###  UseDistanceShore covariate and tensor splines ti

#### Fit sde

@@

crcvm_fjords_hf_ne2<- foreach(i=1:N_BATCH,.packages=c("smoothSDE")) %dopar% {
  #estimation with tensor splines
  formulas <- list(mu1=~1,mu2=~1,tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),omega=~te(theta,DistanceShore,k=SP_DF,bs="cs"))
  par0 <- c(0,0,1,1,0)
  m <- SDE$new(formulas = formulas,data = data_fjords_hf_ne[[paste("BATCH",k,sep="")]],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2"),
                      other_data=list("H"=H_hf))
  m$fit(method="BFGS")
  m 
}

@@

#### Plots estimated smooth in 3D for one batch

```{r}
xmin=list("theta"=-pi,"DistanceShore"=0.3)
xmax=list("theta"=pi,"DistanceShore"=2)
res=crcvm_fjords_hf_ne2$BATCH1$get_all_plots(baseline=NULL,model_name="crcvm_fjords_hf_ne2",xmin=xmin,xmax=xmax,
                                    xlabel=xlabel,link=link,show_CI="pointwise",true_smooths=true_smooths,save=TRUE)
res$fe_omega_theta_DistanceShore
```

#### Violin plot of the estimated spline coefficients for one batch

@@

#### Simulate from fitted model

@@

#### Violin plot of the estimated spline coefficients for all batches

n_coeffs=length(sp_coeff_Dshore)
coeffs_df=data.frame("estimate"=NA,"coeff_name"=NA)

for (k in 1:N_BATCH) {
  m<-crcvm_fjords_hf_ne2[[paste("BATCH",k,sep="")]]
  coeffs=m$coeff_re()
  coeff_names=rownames(coeffs)
  coeff_values=as.numeric(coeffs)
  sub_index=grepl("omega",coeff_names)
  sub_coeffs_df=data.frame("estimate"=coeff_values[sub_index],"coeff_name"=factor(coeff_names[sub_index]))
  coeffs_df=rbind(coeffs_df,sub_coeffs_df)
}

>>>>

# SET UP

## Import libraries

@@

#parallel computing
library(foreach)
library(doParallel)

@@

```{r}
#number of batch
N_BATCH=5

@@

## Estimate with ExpShore covariate and tensor splines te

### Fit sde

@@

crcvm_fjords_hf_ne1<- foreach(k=1:N_BATCH,.packages=c("smoothSDE")) %do% {
  #estimation with tensor splines
  formulas <- list(mu1=~1,mu2=~1,tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),omega=~te(theta,ExpShore,k=SP_DF,bs="cs"))
  par0 <- c(0,0,1,1,0)
  m <- SDE$new(formulas = formulas,data = data_fjords_hf_ne[[paste("BATCH",k,sep="")]],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2"),
                      other_data=list("H"=H_hf))
  m$fit(method="BFGS")
  m 
}

@@

### Plot estimated smooth parameter omega in 3D for one batch

@@

### Violin plot of the estimated spline coefficients for one batch

```{r violin plots spline coeff te ExpShore for one batch of low error samples}

@@

### Violin plot of the estimated spline coefficients for all batches

@@

  #fitted model
  m<-crcvm_fjords_hf_ne1[[paste("BATCH",k,sep="")]]

  #re and fe coeffs
  coeffs=rbind(m$coeff_re(),m$coeff_fe())

  #names and values of coeffs
  coeff_names=rownames(coeffs)
  coeff_values=as.numeric(coeffs)

  #keep coeffs that do not contain ID
  sub_index=grep("ID",coeff_names,invert=TRUE)

  #standard deviation of random effects
  sdev=m$sdev()
  sdev_names=rownames(sdev)[1:2]
  sdev_values=as.numeric(sdev)[1:2]

  #bind dataframe
  sub_coeffs_df=data.frame("estimate"=c(coeff_values[sub_index],sdev_values),"coeff_name"=factor(c(coeff_names[sub_index],sdev_names)))
  coeffs_df=rbind(coeffs_df,sub_coeffs_df)
}

@@

p=ggplotly(p)
saveWidget(p, file ="crcvm_fjords_hf_ne1/all_batches_coeffs_crcvm_fjords_hf_ne1.html")

p
```

### Biais and RMSE

```{r compute biais and rmse ExpShore splines}
biais=c()
rmse=c()

#biais and rmse for omega coefficients
for (string in names(sp_coeff_ExpShore)) {
  name=paste("omega.",string,sep="")
  true_value=sp_coeff_ExpShore[string]
  estimates=coeffs_df[coeffs_df$coeff_name==name,"estimate"]
  biais=c(biais,1/N_BATCH*sum(abs(estimates-true_value)))
  rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_value)^2)))
}

#add biais and rmse for tau
true_log_tau=0
estimates=coeffs_df[coeffs_df$coeff_name=="tau.(Intercept)","estimate"]
biais=c(biais,1/N_BATCH*sum(abs(estimates-true_log_tau)))
rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_log_tau)^2)))

#add biais and rmse for nu
true_log_nu=log(4)
estimates=coeffs_df[coeffs_df$coeff_name=="nu.(Intercept)","estimate"]
biais=c(biais,1/N_BATCH*sum(abs(estimates-true_log_nu)))
rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_log_nu)^2)))

#add biais and rmse for sdev tau
true_sdev_tau=0.1
estimates=coeffs_df[coeffs_df$coeff_name=="tau.(ID)","estimate"]
biais=c(biais,1/N_BATCH*sum(abs(estimates-true_sdev_tau)))
rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_sdev_tau)^2)))

#add biais and rmse for sdev tau
true_sdev_nu=0.3
estimates=coeffs_df[coeffs_df$coeff_name=="nu.(ID)","estimate"]
biais=c(biais,1/N_BATCH*sum(abs(estimates-true_sdev_nu)))
rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_sdev_nu)^2)))

names(biais)=c(paste("omega.",names(sp_coeff_ExpShore),sep=""),"tau.(Intercept)","nu.(Intercept)","tau.(ID)","nu.(ID)")
names(rmse)=c(paste("omega.",names(sp_coeff_ExpShore),sep=""),"tau.(Intercept)","nu.(Intercept)","tau.(ID)","nu.(ID)")

@@

```{r print biais ExpShore splines}
biais
```

```{r print rmse ExpShore splines}
rmse
```

### Simulate from fitted model

@@

### Plot simulation from fitted model

@@

##  Estimate with DistanceShore covariate and tensor splines te

### Fit sde

@@

crcvm_fjords_hf_ne2<- foreach(k=1:N_BATCH,.packages=c("smoothSDE")) %do% {
  #estimation with tensor splines
  formulas <- list(mu1=~1,mu2=~1,tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),omega=~te(theta,DistanceShore,k=SP_DF,bs="cs"))
  par0 <- c(0,0,1,1,0)
  m <- SDE$new(formulas = formulas,data = data_fjords_hf_ne[[paste("BATCH",k,sep="")]],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2"),
                      other_data=list("H"=H_hf))
  m$fit(method="BFGS")
  m 
}

@@

### Plots estimated smooth in 3D for one batch

```{r get plots from hf e estimates with DistanceShore}
xmin=list("theta"=-pi,"DistanceShore"=0.3)
xmax=list("theta"=pi,"DistanceShore"=2)
res=crcvm_fjords_hf_ne2$BATCH1$get_all_plots(baseline=NULL,model_name="crcvm_fjords_hf_ne2",xmin=xmin,xmax=xmax,
                                    xlabel=xlabel,link=link,show_CI="pointwise",true_smooths=true_smooths,save=TRUE)
res$fe_omega_theta_DistanceShore
```

### Violin plot of the estimated spline coefficients for one batch

```{r violin plots spline coeff te DistanceShore for one batch of low error samples}

@@

### Violin plot of the estimated spline coefficients for all batches

n_coeffs=length(sp_coeff_DistanceShore)
coeffs_df=data.frame("estimate"=NA,"coeff_name"=NA)

  #fitted model
  m<-crcvm_fjords_hf_ne2[[paste("BATCH",k,sep="")]]

  #re and fe coeffs
  coeffs=rbind(m$coeff_re(),m$coeff_fe())

  #names and values of coeffs
  coeff_names=rownames(coeffs)
  coeff_values=as.numeric(coeffs)

  #keep coeffs that do not contain ID
  sub_index=grep("ID",coeff_names,invert=TRUE)

  #standard deviation of random effects
  sdev=m$sdev()
  sdev_names=rownames(sdev)[1:2]
  sdev_values=as.numeric(sdev)[1:2]

  #bind dataframe
  sub_coeffs_df=data.frame("estimate"=c(coeff_values[sub_index],sdev_values),"coeff_name"=factor(c(coeff_names[sub_index],sdev_names)))
  coeffs_df=rbind(coeffs_df,sub_coeffs_df)
}

p
```

### Biais and RMSE

```{r compute biais and rmse DistanceShore splines}
biais=c()
rmse=c()

#biais and rmse for omega coefficients
for (string in names(sp_coeff_ExpShore)) {
  name=paste("omega.",string,sep="")
  true_value=sp_coeff_ExpShore[string]
  estimates=coeffs_df[coeffs_df$coeff_name==name,"estimate"]
  biais=c(biais,1/N_BATCH*sum(abs(estimates-true_value)))
  rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_value)^2)))
}

#add biais and rmse for tau
true_log_tau=0
estimates=coeffs_df[coeffs_df$coeff_name=="tau.(Intercept)","estimate"]
biais=c(biais,1/N_BATCH*sum(abs(estimates-true_log_tau)))
rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_log_tau)^2)))

#add biais and rmse for nu
true_log_nu=log(4)
estimates=coeffs_df[coeffs_df$coeff_name=="nu.(Intercept)","estimate"]
biais=c(biais,1/N_BATCH*sum(abs(estimates-true_log_nu)))
rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_log_nu)^2)))

#add biais and rmse for sdev tau
true_sdev_tau=0.1
estimates=coeffs_df[coeffs_df$coeff_name=="tau.(ID)","estimate"]
biais=c(biais,1/N_BATCH*sum(abs(estimates-true_sdev_tau)))
rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_sdev_tau)^2)))

#add biais and rmse for sdev tau
true_sdev_nu=0.3
estimates=coeffs_df[coeffs_df$coeff_name=="nu.(ID)","estimate"]
biais=c(biais,1/N_BATCH*sum(abs(estimates-true_sdev_nu)))
rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_sdev_nu)^2)))

names(biais)=c(paste("omega.",names(sp_coeff_ExpShore),sep=""),"tau.(Intercept)","nu.(Intercept)","tau.(ID)","nu.(ID)")
names(rmse)=c(paste("omega.",names(sp_coeff_ExpShore),sep=""),"tau.(Intercept)","nu.(Intercept)","tau.(ID)","nu.(ID)")

```{r print biais DistanceShore splines}
biais
```

```{r print rmse DistanceShore splines}
rmse
```

### Simulate from fitted model

@@

### 2024-05-02 R/simulation_study/simulation_study_rect.Rmd
improve simulation study

### plot one sample and one batch

@@

## Estimate from high frequency data with low error

### Use ExpShore covariate and tensor splines te

#### Fit sde

@@

#### Violin plot of the estimated coefficients for one batch

@@

for (k in 1:N_BATCH) {
  m<-crcvm_rect_hf_ne1[[paste("BATCH",k,sep="")]]
  coeffs=rbind(m$coeff_re(),m$coeff_fe())
  coeff_names=rownames(coeffs)
  coeff_values=as.numeric(coeffs)
  sub_index=grepl("omega",coeff_names)
  sub_coeffs_df=data.frame("estimate"=coeff_values[sub_index],"coeff_name"=factor(coeff_names[sub_index]))
  coeffs_df=rbind(coeffs_df,sub_coeffs_df)
}

@@

```{r compute biais and rmse}
biais=c()
rmse=c()
for (string in names(sp_coeff_ExpShore)) {
  name=paste("omega.",string,sep="")
  true_value=sp_coeff_ExpShore[string]
  estimates=coeffs_df[coeffs_df$coeff_name==name,"estimate"]
  biais=c(biais,1/N_BATCH*sum(abs(estimates-true_value))/true_value)
  rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_value)^2)))
}

names(biais)=paste("omega.",names(sp_coeff_ExpShore),sep="")
names(rmse)=paste("omega.",names(sp_coeff_ExpShore),sep="")
```

```{r print biais}
biais
```

```{r print rmse}
rmse
```

#### Simulate from fitted model

@@

#### Plot simulation from fitted model

@@

###  Use DistanceShore covariate and tensor splines te

#### Fit sde

@@

#### Plots estimated smooth in 3D

@@

#### Violin plot of the estimated spline coefficients

@@

p=ggplotly(p)
saveWidget(p, file ="crcvm_rect_hf_ne2/all_batches_coeffs_crcvm_rect_hf_ne1.html")

@@

#### Simulate from fitted model

>>>>

### plot one batch

@@

## Estimate with ExpShore covariate and tensor splines te

### Fit sde

@@

### Violin plot of the estimated spline coefficients for one batch

@@

  #fitted model
  m<-crcvm_rect_hf_ne1[[paste("BATCH",k,sep="")]]

@@

```{r compute biais and rmse ExpShore splines}
biais=c()
rmse=c()

#biais and rmse for omega coefficients
for (string in names(sp_coeff_ExpShore)) {
  name=paste("omega.",string,sep="")
  true_value=sp_coeff_ExpShore[string]
  estimates=coeffs_df[coeffs_df$coeff_name==name,"estimate"]
  biais=c(biais,1/N_BATCH*sum(abs(estimates-true_value)))
  rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_value)^2)))
}

names(biais)=c(paste("omega.",names(sp_coeff_ExpShore),sep=""),"tau.(Intercept)","nu.(Intercept)","tau.(ID)","nu.(ID)")
names(rmse)=c(paste("omega.",names(sp_coeff_ExpShore),sep=""),"tau.(Intercept)","nu.(Intercept)","tau.(ID)","nu.(ID)")

```{r print biais ExpShore splines}
biais
```

```{r print rmse ExpShore splines }
rmse
```

### Simulate from fitted model

@@

### Plot simulation from fitted model

@@

## Estimate with ExpShore covariate and tensor splines te

### Fit sde

@@

### Plots estimated smooth in 3D

@@

### Violin plot of the estimated spline coefficients

@@

p=ggplotly(p)
saveWidget(p, file ="crcvm_rect_hf_ne2/all_batches_coeffs_crcvm_rect_hf_ne2.html")

@@

### Simulate from fitted model
